---
title: "PopSongs"
output: html_document
date: "2026-04-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# ===========================================================
# 1: POP SETUP
# Purpose:
# - Load all required packages
# - Load pop songs dataset
# - Store Genius API token
# Inputs:
# - Pop_Songs1.csv
# Outputs:
# - pop_songs dataframe
# ===========================================================

```{r pop_setup, message=FALSE, warning=FALSE}

#Loaded packages needed for:
library(rvest)    #Web Scraping
library(stringr)  #String Manipluation
library(dplyr)    #Data Manipulation
library(purrr)    #Function Programing
library(tibble)   #Data framing tools
library(httr)     #API requests
library(jsonlite) #JSON tools
library(readr)    #Reading CSV Files
library(tidytext) #Stop word dataset
library(ggplot2)  #Visualization

#Genius API token used for Genius song searching
#Kept this unique variable name so it does not conflict with my group members' code.
GENIUS_TOKEN_POP <- "nkuSlxBMD3TMJ40tfIEKNMgAi__zgTH2bV9vAAu0OqtEHXhOchXP2f_xIljeL1Bi"

#Read in pop songs dataset
pop_songs <- read_csv("/Users/kyleedahlstrom/Downloads/Pop_Songs1.csv")

#Preview the dataset
head(pop_songs)

```

# ===========================================================
# 2: CLEAN TEXT FOR GENIUS SEARCH
# Purpose:
# - Clean artist names and song titles before searching
# - Remove special characters and formatting issues
# Inputs:
# - Artist names
# - Song titles
# Outputs:
# - Cleaned text strings
# ===========================================================

```{r pop_text_cleaning}

pop_clean_text <- function(x) {
  
  x %>%
    #Convert to character format
    as.character() %>%
    #Fix encoding issues
    enc2utf8() %>%
    #Convert special characters into standard text
    iconv(from = "UTF-8", to = "ASCII//TRANSLIT", sub = "") %>%
    #Remove non-alphanumeric characters
    str_replace_all("[^[:alnum:]\\s']", " ") %>%
    #Remove extra spaces
    str_squish()
}
```

# ===========================================================
# 3: SEARCH GENIUS FOR SONG URL
# Purpose:
# - Search Genius API
# - Find best matching Genius page for each song
# Inputs:
# - Artist name
# - Song title
# Outputs:
# - Genius song URL
# ===========================================================

```{r pop_genius_search}

pop_search_genius_url <- function(artist, title, token = GENIUS_TOKEN_POP) {
  
  #Combine artist + title into one search query
  pop_query <- paste(
    pop_clean_text(artist),
    pop_clean_text(title))
  
  #Send request to Genius API
  pop_res <- GET(
    "https://api.genius.com/search",
    query = list(q = pop_query),
    add_headers(
      Authorization = paste("Bearer", token)))
  
  #If request fails, return NA
  if (status_code(pop_res) != 200)
    return(NA_character_)
  
  #Convert JSON response into R object
  pop_dat <- fromJSON(
    content(pop_res, as = "text", encoding = "UTF-8"), flatten = TRUE)
  
  #Extract search results
  pop_hits <- pop_dat$response$hits
  
  #If no results found, return NA
  if (length(pop_hits) == 0 || nrow(pop_hits) == 0)
    return(NA_character_)
  
  #Return first Genius result URL
  pop_hits$result.url[1]
}
```

# ===========================================================
# 4: SCRAPE LYRICS FROM GENIUS URL
# Purpose:
# - Visit Genius webpage
# - Extract lyrics text
# Inputs:
# - Genius URL
# Outputs:
# - Lyrics text
# ===========================================================

```{r pop_genius_scraping}

pop_get_lyrics_from_url <- function(url) {
  
  #If URL missing, return NA
  if (is.na(url) || url == "")
    return(NA_character_)
  
  #Pause between requests to prevent Genius from blocking scraping
  Sys.sleep(runif(1, 3, 6))
  
  #Send webpage request
  #User agent makes request appear more like a real browser
  pop_res <- tryCatch(
    
    GET(url, user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64)"),
      add_headers(`Accept-Language` = "en-US,en;q=0.9",
        `Accept` =
          "text/html,application/xhtml+xml,application/xml;q=0.9,*/*;q=0.8")),
    
    error = function(e) return(NULL))
  
  #If request failed, return NA
  if (is.null(pop_res) || status_code(pop_res) != 200)
    return(NA_character_)
  
  #Read HTML page
  pop_page <- tryCatch(
    read_html(pop_res),
    error = function(e) return(NULL))
  
  #If page fails, return NA
  if (is.null(pop_page))
    return(NA_character_)
  
  #Extract lyrics containers
  pop_lyrics <- pop_page %>%
    
    html_elements("div[data-lyrics-container='true']") %>%
    
    html_text2() %>%
    
    paste(collapse = "\n")
  
  #If no lyrics found, return NA
  if (pop_lyrics == "" || length(pop_lyrics) == 0)
    return(NA_character_)
  
  #Clean ending "Embed" text
  pop_lyrics %>%
    str_replace_all("Embed$", "") %>%
    str_squish()
}

```

# ===========================================================
# 5: COLLECT LYRICS
# Purpose:
# - Search Genius for each song
# - Scrape lyrics
# - Save progress continuously
# Inputs:
# - pop_songs dataframe
# Outputs:
# - Pop_songs_with_lyrics.csv
# ===========================================================

```{r pop_collect_lyrics, eval=FALSE}

pop_songs_with_lyrics <- pop_songs %>%
  
#Create simplified artist name
#This improves Genius search matching
#Find Genius page URL
  mutate(
    
    main_artist =
      str_split(
        Artist,
        ",|&|feat\\.|featuring"
      ) %>%
      map_chr(1),
    
    genius_url =
      map2_chr(
        main_artist,
        Song,
        pop_search_genius_url
      )
  )

#Create empty vector to store lyrics
pop_lyrics_list <- vector(
  "character",
  nrow(pop_songs_with_lyrics)
)

#Loop through songs one-by-one
#More reliable than map()
for (i in seq_len(nrow(pop_songs_with_lyrics))) {
  
  #Print progress in console
  message(
    "Pulling lyrics for song ",
    i,
    " of ",
    nrow(pop_songs_with_lyrics),
    ": ",
    pop_songs_with_lyrics$Song[i]
  )
  
  #Scrape lyrics
  pop_lyrics_list[i] <-
    pop_get_lyrics_from_url(
      pop_songs_with_lyrics$genius_url[i]
    )
  
  #Save lyrics into dataset
  pop_songs_with_lyrics$lyrics <- pop_lyrics_list
  
  #Save progress file after EVERY song
  #Prevents losing progress if code stops
  write.csv(
    pop_songs_with_lyrics,
    "Pop_songs_with_lyrics_PROGRESS.csv",
    row.names = FALSE
  )
}

#Save final lyrics dataset
write.csv(
  pop_songs_with_lyrics,
  "Pop_songs_with_lyrics.csv",
  row.names = FALSE
)

#Count successful lyric pulls
sum(!is.na(pop_songs_with_lyrics$lyrics))

#Count failed lyric pulls
sum(is.na(pop_songs_with_lyrics$lyrics))

```

# ===========================================================
# 6: LOAD LYRICS DATASET
# Purpose:
# - Load dataset that already contains lyrics
# Inputs:
# - Pop_songs_with_lyrics.csv
# Outputs:
# - pop_songs_with_lyrics dataframe
# ===========================================================

```{r pop_lyrics_dataset, message=FALSE, warning=FALSE}

#Read in New CSV
pop_songs_with_lyrics <-read_csv("Pop_songs_with_lyrics.csv")

head(pop_songs_with_lyrics)

```

# ===========================================================
# 7: CLEAN LYRICS
# Purpose:
# - Remove Genius metadata
# - Remove section labels
# - Remove junk text
# Inputs:
# - Raw lyrics
# Outputs:
# - Cleaned lyrics
# ===========================================================

```{r pop_clean_lyrics}

pop_clean_lyrics <- function(lyrics) {
  
  #Return NA if lyrics missing
  if (is.na(lyrics) || lyrics == "")
    return(NA_character_)
  
  #Common metadata words
  metadata_words <- regex(
    
    "written|recorded|released|album|single|billboard|chart|song by|music video",
    
    ignore_case = TRUE
  )
  
  pop_cleaned <- lyrics %>%
    
    #Remove Everything before "Read More" to get rid of captions
    str_replace("^.*?Read More", "") %>%
    #Remove text before "Lyrics"
    str_replace("^.*?Lyrics", "") %>%
    #Remove [Verse], [Chorus], etc.
    str_replace_all("\\[.*?\\]", " ") %>%
    #Remove contributor info
    str_replace_all("\\d+ Contributors?", " ") %>%
    #Remove "Embed"
    str_replace_all("\\d*Embed$", " ") %>%
    #Remove extra spaces
    str_squish()
  
  #Remove leftover metadata sentences
  for (i in 1:5) {
    
    first_sentence <-
      str_extract(pop_cleaned, "^.*?\\.")
    
    if (!is.na(first_sentence) &&
        str_detect(first_sentence,
                   metadata_words)) {
      
      pop_cleaned <- pop_cleaned %>%
        str_remove("^.*?\\.") %>%
        str_squish()
    }
  }
  
  pop_cleaned
}
```

# ===========================================================
# 8: CLEAN POP DATASET
# Purpose:
# - Apply lyric cleaning
# - Create word count variable
# Inputs:
# - pop_songs_with_lyrics dataframe
# Outputs:
# - pop_songs_clean dataframe
# ===========================================================

```{r pop_cleaned_dataset}

#Preview cleaned dataset
pop_songs_clean <- pop_songs_with_lyrics %>%
  
  #Clean lyrics
  mutate(lyrics = map_chr(lyrics, pop_clean_lyrics)) %>%
  
  #Remove missing lyrics
  filter(!is.na(lyrics), lyrics != "") %>%
  
  #Remove duplicate songs
  distinct(Song, .keep_all = TRUE)

#Save cleaned dataset
write.csv(
  pop_songs_clean,
  "pop_songs_clean.csv",
  row.names = FALSE
)

#Preview cleaned dataset
head(pop_songs_clean)

```

# ===========================================================
# 9: LOAD STOP WORDS
# Purpose:
# - Load common English stop words
# - Remove words like: "the", "and", "is", "a"
# Inputs:
# - tidytext stop_words dataset
# Outputs:
# - stop_words dataframe
# ===========================================================

```{r pop_stop_words}

#Load built-in stop words dataset
data("stop_words")
```

# ===========================================================
# 10: WORD REPETITION FUNCTION
# Purpose:
# - Measure how repetitive each pop song is
# - Uses meaningful words only
# Inputs:
# - lyrics text
# Outputs:
# - repetition score
# ===========================================================

```{r pop_word_repetition}

pop_word_repetition_score <- function(lyrics) {
  
  #Return NA if lyrics missing
  if (is.na(lyrics) || lyrics == "")
    return(NA_real_)

  #Convert lyrics to lowercase
  #Remove punctuation
  #Split into words
  pop_words <- lyrics %>%
    
    tolower() %>%
    
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    
    str_split("\\s+") %>%
    
    unlist()

  #Remove blank entries
  pop_words <- pop_words[pop_words != ""]

  #Remove stop words
  pop_words <- pop_words[
    !pop_words %in% stop_words$word
  ]

  #If no meaningful words remain
  #return NA
  if (length(pop_words) == 0)
    return(NA_real_)

  # Repetition score formula: 1 - (unique words / total words)
  #Higher value = more repetition
  1 - (length(unique(pop_words)) / length(pop_words))
}
```

# ===========================================================
# 11: WORD COUNT FUNCTION
# Purpose:
# - Count number of words in each pop song
# Inputs:
# - lyrics text
# Outputs:
# - total word count
# ===========================================================

```{r pop_word_count}

pop_word_count <- function(lyrics) {
  
  #Return NA if lyrics missing
  if (is.na(lyrics) || lyrics == "")
    return(NA_real_)

  #Clean and split lyrics into words
  pop_words <- lyrics %>%
    
    tolower() %>%
    
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    
    str_split("\\s+") %>%
    
    unlist()

  # Remove blank words
  pop_words <- pop_words[pop_words != ""]

  # Return total number of words
  length(pop_words)
  
  #Add word count variable to dataset
pop_songs_clean <- pop_songs_clean %>%
  
  mutate(pop_number_of_words =
           map_dbl(lyrics, pop_word_count))
}
```

# ===========================================================
# 12: ADD WORD REPETITION TO DATASET
# Purpose:
# - Calculate repetition score for each pop song
# Inputs:
# - pop_songs_clean dataframe
# Outputs:
# - Updated dataframe with: pop_word_repeat_score
# ===========================================================

```{r pop_updated_datset}

pop_songs_clean <- pop_songs_clean %>%
  
  #Calculate repetition score
  mutate(pop_word_repeat_score = map_dbl(lyrics, pop_word_repetition_score))

#Preview results
pop_songs_clean %>%
  select(Song, pop_word_repeat_score) %>%
  
  head()
```

# ===========================================================
# 13: TITLE REPETITION FUNCTION
# Purpose:
# - Count how often song title appears in lyrics
# Inputs:
# - song title
# - lyrics text
# Outputs:
# - title repetition count
# ===========================================================

```{r pop_title_repetition}

pop_title_repetition_score <- function(pop_title, pop_lyrics) {
  
  #Return NA if title or lyrics missing
  if (is.na(pop_title) ||
      is.na(pop_lyrics) ||
      pop_lyrics == "")
    
    return(NA_real_)

  #Clean title text
  pop_title_clean <- pop_title %>%
    
    tolower() %>%
    
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    
    str_squish()

  #Clean lyrics text
  pop_lyrics_clean <- pop_lyrics %>%
    
    tolower() %>%
    
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    
    str_squish()

  #Count how many times title appears
  str_count(pop_lyrics_clean, fixed(pop_title_clean))
}
```

# ===========================================================
# 14: ADD TITLE REPETITION TO DATASET
# Purpose:
# - Calculate title repetition score
# Inputs:
# - pop_songs_clean dataframe
# Outputs:
# - Updated dataframe with: pop_title_repeat_score
# ===========================================================

```{r pop_updated_dataset}

pop_songs_clean <- pop_songs_clean %>%
  
  #Calculate title repetition count
  mutate(pop_title_repeat_score = map2_dbl(
        Song, lyrics, pop_title_repetition_score))

#Preview results
pop_songs_clean %>%
  
  select(Song, pop_title_repeat_score) %>%
  
  head()
```

# ===========================================================
# 15: YEARLY SUMMARY
# Purpose:
# - Calculate yearly averages
# Inputs:
# - repetition variables
# Outputs:
# - pop_year_summary dataframe
# ===========================================================

```{r pop_yearly_summary}

pop_year_summary <- pop_songs_clean %>%
  
  group_by(Year) %>%
  
  #Average word repetition score
  summarise(pop_avg_word_repeat = mean(pop_word_repeat_score, na.rm = TRUE),
    
    #Average title repetition count
    pop_avg_title_repeat = mean(
      pop_title_repeat_score, na.rm = TRUE), .groups = "drop") %>%
  
  #Remove invalid rows
  filter(!is.na(pop_avg_word_repeat), !is.nan(pop_avg_word_repeat))

#Preview summary
pop_year_summary
```

# ===========================================================
# 16: WORD REPETITION PLOT
# Purpose:
# - Visualize repetition trends over time
# Inputs:
# - pop_year_summary dataframe
# Outputs:
# - line graph
# ===========================================================

```{r pop_word_repetition_plot, warning=FALSE, message=FALSE}

ggplot(pop_year_summary, aes(
  x = Year, y = pop_avg_word_repeat)) +
  
  #Trend line
  geom_line(linewidth = 1) +
  
  #Data points
  geom_point(size = 2) +
  
  #Smooth trend line
  geom_smooth(color = "blue", se = FALSE, linetype = "dashed") +
  
  #Clean theme
  theme_minimal() +
  
  #Improve readability
  theme(plot.title = element_text(
      size = 16, face = "bold", hjust = 0.5),
    
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)) +
  
  labs(title =
      "Word Repetition in Pop Songs Over Time (2006–2025)",
      subtitle =
      "Average repetition score per year",
      x = "Year", y = "Repetition Score")
```

The results show how repetitive pop song lyrics are over time. Higher repetition scores indicate that songs reuse meaningful words more frequently throughout the lyrics. The trend line helps show whether repetition is increasing or decreasing across years. Changes in repetition may reflect shifts in songwriting style, choruses, or lyrical simplicity in pop music. The repetitiveness of pop songs seems to hit its peak in the year 2014, then gradually decrease with 2025 being one of the lowest repetition scores.

# ===========================================================
# 17: TITLE REPETITION PLOT
# Purpose:
# - Visualize title repetition trends
# Inputs:
# - pop_year_summary dataframe
# Outputs:
# - line graph
# ===========================================================

```{r , warning=FALSE, message=FALSE}

ggplot(pop_year_summary, aes(x = Year, y = pop_avg_title_repeat)) +
  
  #Main line
  geom_line(linewidth = 1) +
  
  #Points
  geom_point(size = 2) +
  
  #Smooth trend line
  geom_smooth(color = "blue", se = FALSE, linetype = "dashed") +
  
  #Minimal theme
  theme_minimal() +
  
  #Better readability
  theme(plot.title = element_text(
      size = 16, face = "bold", hjust = 0.5),
    
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)) +
  
  labs(title =
      "Title Repetition in Pop Songs Over Time (2006–2025)",
      subtitle =
      "Average number of times song titles appear in lyrics",
      x = "Year", y = "Average Title Repetitions")
```

The title repetition results show how often pop songs repeat their titles within the lyrics. Songs with highly repetitive choruses or hooks often repeat the title many times. This graph helps identify whether title repetition has become more common in modern pop music. For pop music, it looks like the title repetition has gone down over the more recent years. 


# ===========================================================
# 18: TOP WORDS OVERALL
# Purpose:
# - Find most common meaningful words
# - Remove filler/interjection words
# Inputs:
# - pop_songs_clean dataframe
# Outputs:
# - pop_top_words dataframe
# ===========================================================

```{r pop_top_words, message=FALSE, warning=FALSE}

#Words to Remove from analysis
pop_filler_words <- c(
  "ayy", "ay", "uh", "ah", "ha", "hah",
  "woo", "woah", "whoa", "la", "doh",
  "mm", "mmm", "hmm", "ohh", "yah",
  "ya", "eh", "ow", "hey", "yo", "em", "da"
)

pop_top_words <- pop_songs_clean %>%
  
  #Split lyrics into individual words
  unnest_tokens(word, lyrics) %>%
  
  #Remove standard stop words
  anti_join(stop_words, by = "word") %>%
  
  #Remove filler/interjection words
  filter(!word %in% pop_filler_words) %>%
  
  #Count remaining words
  count(word, sort = TRUE)

#View top 20 words
pop_top_words %>%
  head(20)
```

# ===========================================================
# 19: MOST REPETITIVE POP SONGS
# Purpose:
# - Find songs with highest repetition scores
# Inputs:
# - pop_songs_clean dataframe
# Outputs:
# - table of repetitive songs
# ===========================================================

```{r pop_most_repetitive}

pop_songs_clean %>%
  
  arrange(desc(pop_word_repeat_score)) %>%
  
  select(Song, Artist, pop_word_repeat_score) %>%
  
  head(10)
```

