---
title: "STAT 345 Final Project"
author: "Melia, Gracie, Kylee
output: word_document
date: "2026-04-15"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

### Rap Songs Analysis

## Rap Setup

```{r rap_setup, message=FALSE, warning=FALSE}
## Load required packages and dataset
# Purpose: Ensure all packages needed are available and dataset (with lyrics) is loaded
# Inputs: rap_songs_with_lyrics.csv (originally rap_songs.csv)
# Outputs: rap_songs dataframe

# Load libraries for scraping, cleaning, and analysis
library(rvest)      # Web scraping
library(stringr)    # String manipulation
library(dplyr)      # Data manipulation
library(purrr)      # Functional programming
library(tibble)     # Data frame tools
library(httr)       # API requests
library(jsonlite)   # JSON parsing
library(ggplot2)    # Visualization
library(tidytext)   # Stop word dataset

# Genius API token (I removed my personal token)
GENIUS_TOKEN <- "Token"

# Read in dataset for analysis
# NOTE: Original dataset was "rap_songs.csv" (manually created with song info).
# After collecting lyrics using the Genius API and expanding the dataset to include them, I saved the dataset as "rap_songs_with_lyrics.csv", and used this for all analysis steps.
rap_songs <- read.csv("rap_songs_with_lyrics.csv", stringsAsFactors = FALSE)
```


## Text Cleaning for Genius Search

```{r rap_text_cleaning}
## Clean text for Genius API search
# Purpose: Standardize artist and song title text before searching Genius
# Inputs: Character string
# Outputs: Cleaned text string

clean_text <- function(x) {
  x %>%
    enc2utf8() %>%
    iconv(from = "UTF-8", to = "ASCII//TRANSLIT", sub = "") %>%
    str_replace_all("[“”]", "\"") %>%
    str_replace_all("[‘’]", "'") %>%
    str_squish()
}

```


## Genius API Search

```{r rap_genius_search}
## Search Genius for song URL
# Purpose: Find the best matching Genius lyrics page for each song
# Inputs: Artist name, song title, and Genius API token
# Outputs: Genius URL for the best matching result

search_genius_url <- function(main_artist, title, token = GENIUS_TOKEN) {
  # Create search query using cleaned artist and title
  query <- paste(clean_text(main_artist), clean_text(title))

  # Send request to Genius API
  res <- GET(
    url = "https://api.genius.com/search",
    query = list(q = query),
    add_headers(Authorization = paste("Bearer", token))
  )

  # Return missing value if request is unsuccessful
  if (status_code(res) != 200) return(NA_character_)

  # Convert API response from JSON into an R object
  txt <- content(res, as = "text", encoding = "UTF-8")
  dat <- fromJSON(txt, flatten = TRUE)

  # Extract search results
  hits <- dat$response$hits
  # Return missing value if no Genius results are found
  if (length(hits) == 0 || nrow(hits) == 0) return(NA_character_)

  # Clean artist and title for comparison
  artist_clean <- tolower(clean_text(main_artist))
  title_clean  <- tolower(clean_text(title))

  # Score results based on how closely they match artist and title
  result_df <- tibble(
    url = hits$result.url,
    artist = tolower(clean_text(hits$result.primary_artist.name)),
    song = tolower(clean_text(hits$result.title))
  ) %>%
    mutate(
      artist_match = artist == artist_clean,
      song_match = song == title_clean,
      score = 2 * artist_match + 2 * song_match +
              str_detect(song, fixed(title_clean)) +
              str_detect(artist, fixed(artist_clean))
    ) %>%
    arrange(desc(score))

  # Return the highest-scoring Genius URL
  result_df$url[1]
}

```


## Genius Lyric Scraping

```{r rap_genius_scraping}
## Scrape lyrics from Genius URL
# Purpose: Extract lyrics text from the selected Genius lyrics page
# Inputs: Genius song URL
# Outputs: Raw lyrics text

get_lyrics_from_genius_url <- function(song_url) {
  # Return missing value if URL is unavailable
  if (is.na(song_url) || song_url == "") return(NA_character_)

  # Try to read webpage and return NULL if page cannot be read
  page <- tryCatch(
    read_html(song_url),
    error = function(e) NULL
  )

  # Return missing value if webpage was not read successfully
  if (is.null(page)) return(NA_character_)

  # Select lyrics containers from Genius page
  nodes <- html_elements(page, "[data-lyrics-container='true']")
  if (length(nodes) == 0) return(NA_character_)

  # Combine lyrics text into one string
  lyrics <- nodes %>%
    html_text2() %>%
    paste(collapse = "\n") %>%
    str_replace_all("\\n{2,}", "\n") %>%
    str_trim()

  # Find likely beginning of lyrics section
  start_pos <- str_locate(lyrics, "\\[(Intro|Verse|Chorus|Hook|Bridge|Outro)")[1]

  # Remove extra text before lyrics begin, if found
  if (!is.na(start_pos)) {
    lyrics <- str_sub(lyrics, start_pos)
  }

  lyrics <- str_trim(lyrics)

  # Return missing value if lyrics are empty
  if (lyrics == "") return(NA_character_)

  lyrics
}

```


## Combined Lyric Retrieval

```{r rap_combined_lyrics_retrieval}
## Retrieve lyrics using Genius API
# Purpose: Combine Genius URL search and lyric scraping into one function
# Inputs: Artist name, song title, and Genius API token
# Outputs: Lyrics text for each song

get_lyrics_genius_api <- function(main_artist, title, token = GENIUS_TOKEN) {
  
  # Search for the Genius URL for the given artist and title
  url <- search_genius_url(main_artist, title, token)
  
  # Pause briefly to avoid sending too many API requests at once
  Sys.sleep(0.5)
  
  # Scrape and return lyrics from the Genius URL
  get_lyrics_from_genius_url(url)
}

```


## Lyric Cleaning

```{r rap_lyric_cleaning}
## Clean scraped lyrics
# Purpose: Remove section labels (like [chorus 1]) and clean formatting in lyrics
# Inputs: Raw lyrics text
# Outputs: Cleaned lyrics text

clean_lyrics_text <- function(lyrics) {
  
  # Return NA if lyrics are missing
  if (is.na(lyrics) || lyrics == "") return(NA_character_)

  lyrics <- lyrics %>%
    str_replace_all("\\[.*?\\]", "") %>%  # Remove [Verse], [Chorus], etc.
    str_replace_all("\r", "\n") %>%
    str_replace_all("\\n{2,}", "\n") %>%
    str_replace_all("([a-z])([A-Z])", "\\1 \\2") %>%
    str_replace_all("[ \t]+", " ") %>%
    str_trim()

  lyrics
}

```


## Lyric Collection

```{r rap_lyric_collection, eval=FALSE}
## Collect lyrics and save expanded dataset
# Purpose: Add lyrics to the original song dataset
# Inputs: rap_songs dataframe with artist and title columns
# Outputs: rap_songs_with_lyrics.csv
# NOTE: This chunk was run once to collect lyrics and is not re-run when knitting.

rap_songs <- rap_songs %>%
  mutate(
    # Retrieve lyrics for each artist/title pair
    lyrics = map2_chr(main_artist, title, ~ get_lyrics_genius_api(.x, .y)),
    
    # Clean scraped lyrics text
    lyrics = map_chr(lyrics, clean_lyrics_text)
  )

# Save dataset with lyrics
write.csv(rap_songs, "rap_songs_with_lyrics.csv", row.names = FALSE)

```


## Re-Clean Saved Lyrics

```{r rap_reclean_lyrics, eval=FALSE}
## Re-clean saved lyrics
# Purpose: Apply updated lyric cleaning to the saved lyrics dataset
# Inputs: rap_songs dataframe with existing lyrics column
# Outputs: rap_songs dataframe with cleaned lyrics

rap_songs <- rap_songs %>%
  mutate(
    # Remove section labels and clean formatting from already-collected lyrics
    lyrics = map_chr(lyrics, clean_lyrics_text)
  )

# Save updated lyrics dataset
write.csv(rap_songs, "rap_songs_with_lyrics.csv", row.names = FALSE)

```


## Stop Words

```{r rap_stop_words}
## Load stop words
# Purpose: Load common English stop words to remove from word repetition calculation
# Inputs: tidytext stop_words dataset
# Outputs: stop_words dataframe available for filtering

data("stop_words")

```


## Word Repetition Function

```{r rap_word_repetition_function}
## Calculate word repetition score
# Purpose: Measure how repetitive each song is based on repeated non-stop words
# Inputs: Lyrics text
# Outputs: Numeric repetition score

word_repetition_score <- function(lyrics) {
  
  # Return missing value if lyrics are missing
  if (is.na(lyrics) || lyrics == "") return(NA_real_)

  # Clean lyrics and split into individual words
  words <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_split("\\s+") %>%
    unlist()

  # Remove blank entries
  words <- words[words != ""]

  # Remove common stop words such as "the", "and", and "to"
  words <- words[!words %in% stop_words$word]

  # Return missing value if no meaningful words remain
  if (length(words) == 0) return(NA_real_)

  # Calculate proportion of repeated words
  1 - (length(unique(words)) / length(words))
}

```


## Word Count Function

```{r rap_word_count_function}
## Count words in each song
# Purpose: Measure song length using total number of words in the lyrics
# Inputs: Lyrics text
# Outputs: Numeric word count

word_count <- function(lyrics) {
  
  # Return missing value if lyrics are missing
  if (is.na(lyrics) || lyrics == "") return(NA_real_)

  # Clean lyrics and split into individual words
  words <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_split("\\s+") %>%
    unlist()

  # Remove blank entries
  words <- words[words != ""]

  # Return total number of words
  length(words)
}

```


## Title Repetition Function

```{r rap_title_repetition_function}
## Calculate title repetition score
# Purpose: Count how many times the song title appears in the lyrics
# Inputs: Song title and lyrics text
# Outputs: Numeric count of title repetitions

title_repetition_score <- function(title, lyrics) {
  
  # Return missing value if title or lyrics are missing
  if (is.na(title) || is.na(lyrics) || lyrics == "") return(NA_real_)

  # Clean title text
  title_clean <- title %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_squish()

  # Clean lyrics text
  lyrics_clean <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_squish()

  # Count number of times title appears in lyrics
  str_count(lyrics_clean, fixed(title_clean))
}

```


## Final Dataset Creation

```{r rap_final_dataset_creation}
## Add repetition scores and word count to dataset
# Purpose: Create final analysis variables for each song
# Inputs: rap_songs dataframe with title and lyrics columns
# Outputs: rap_songs_final.csv

rap_songs <- rap_songs %>%
  mutate(
    # Count total number of words in each song
    number_of_words = map_dbl(lyrics, word_count),
    
    # Calculate repeated word score for each song
    word_repeat_score = map_dbl(lyrics, word_repetition_score),
    
    # Count how many times each song title appears in its lyrics
    title_repeat_score = map2_dbl(title, lyrics, title_repetition_score)
  )

# Save final dataset with analysis variables
write.csv(rap_songs, "rap_songs_final.csv", row.names = FALSE)

```


## Yearly Summary

```{r rap_yearly_summary}
## Summarize repetition by year
# Purpose: Calculate average repetition scores for each year
# Inputs: rap_songs dataframe with repetition score columns
# Outputs: year_summary dataframe for plotting trends over time

year_summary <- rap_songs %>%
  group_by(year) %>%
  summarise(
    # Average word repetition score per year
    avg_word_repeat = mean(word_repeat_score, na.rm = TRUE),
    
    # Average title repetition count per year
    avg_title_repeat = mean(title_repeat_score, na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  filter(!is.na(avg_word_repeat), !is.nan(avg_word_repeat))

```


## Word Repetition Plot

```{r rap_word_repetition_plot, warning=FALSE}
## Plot word repetition over time
# Purpose: Visualize how average word repetition changes by year
# Inputs: year_summary dataframe
# Outputs: Line plot of average word repetition over time

ggplot(year_summary, aes(x = year, y = avg_word_repeat)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_smooth(color = "red", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Word Repetition (Stopwords Removed) in Rap Songs Over Time",
    x = "Year",
    y = "Repetition Score"
  )

```


## Title Repetition Plot

```{r rap_title_repetition_plot, warning=FALSE}
## Plot title repetition over time
# Purpose: Visualize how often song titles are repeated in lyrics by year
# Inputs: year_summary dataframe
# Outputs: Line plot of average title repetition over time

ggplot(year_summary, aes(x = year, y = avg_title_repeat)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_smooth(color = "red", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Title Repetition in Rap Songs Over Time",
    x = "Year",
    y = "Average Number of Title Repetitions"
  )
```


## Most Repetitive Rap Songs

```{r rap_most_repetitive}
## Identify most repetitive rap songs
# Purpose: Display the top 10 songs with the highest word repetition scores
# Inputs: rap_songs dataframe with repetition score variables
# Outputs: Table of the most repetitive songs including year, title, artist, and repetition metrics

rap_songs %>%
  
  # Sort songs from highest to lowest word repetition score
  arrange(desc(word_repeat_score)) %>%
  
  # Select relevant columns for interpretation
  select(year, title, main_artist, word_repeat_score, title_repeat_score, number_of_words) %>%
  
  # Display the top 10 most repetitive songs
  head(10)

```


## Top Words Overall

```{r rap_top_words}
## Find top words overall
# Purpose: Identify most frequently used words in rap lyrics (excluding stop words)
# Inputs: rap_songs dataset with lyrics column
# Outputs: top_words dataframe of most common words

rap_songs <- rap_songs %>%
  mutate(
    lyrics = as.character(lyrics)
  )

top_words <- rap_songs %>%
  
  # Break lyrics into individual words (one word per row)
  unnest_tokens(word, lyrics) %>%
  
  # Remove common stop words (e.g., "the", "and", "to")
  anti_join(stop_words, by = "word") %>%
  
  # Remove offensive word from results
  filter(word != "nigga") %>%
  
  # Count how often each word appears and sort from most to least frequent
  count(word, sort = TRUE)

# Display top 20 most frequent words
top_words %>%
  head(20)


```





--------------------------------

### Pop Songs Analysis





--------------------------------

### Country Songs Analysis

## Country Setup

```{r country_setup, message=FALSE, warning=FALSE}
## Pull lyrics for country songs using the Genius API and Genius song pages
# Purpose: Start with a manually created country songs dataset, locate each song
# on Genius, scrape the lyrics, and save the expanded dataset with lyrics.
# Inputs: country_songs.csv
# Outputs:
#   1. Country_songs_with_lyrics_PROGRESS.csv
#   2. Country_songs_with_lyrics.csv

# Load packages needed for reading files, scraping webpages, cleaning text,
# working with data frames, calling APIs, and reading JSON results.
library(readxl)
library(rvest)
library(stringr)
library(dplyr)
library(purrr)
library(tibble)
library(httr)
library(jsonlite)

# Genius API token used to access the Genius search API.
GENIUS_TOKEN <- "Token"

# Read in the manually created country songs dataset.
# This dataset includes song information such as year, title, and artist.
country_songs <- read_excel("country_songs.csv")

```


## Text Cleaning for Genius Search

```{r country_text_cleaning}
# Function to clean text before searching Genius.
# This helps make artist names and song titles easier to match by:
#   - converting text to character format
#   - fixing encoding issues
#   - removing special characters
#   - removing extra spaces

clean_text <- function(x) {
  x %>%
    as.character() %>%
    enc2utf8() %>%
    iconv(from = "UTF-8", to = "ASCII//TRANSLIT", sub = "") %>%
    str_replace_all("[^[:alnum:]\\s']", " ") %>%
    str_squish()
}

```


## Genius API Search

```{r country_genius_search}
# Function to search Genius for a song and return the best Genius URL.
# Inputs:
#   artist = main artist name
#   title = song title
#   token = Genius API token
# Output:
#   Genius song URL, or NA if no result is found
search_genius_url <- function(artist, title, token = GENIUS_TOKEN) {
  
  # Combine cleaned artist and title into one search query.
  query <- paste(clean_text(artist), clean_text(title))
  
  # Send search request to Genius API.
  res <- GET(
    "https://api.genius.com/search",
    query = list(q = query),
    add_headers(Authorization = paste("Bearer", token))
  )
  
  # If the API request does not work, return NA.
  if (status_code(res) != 200) return(NA_character_)
  
  # Convert the API result from JSON into an R object.
  dat <- fromJSON(content(res, as = "text", encoding = "UTF-8"), flatten = TRUE)
  
  # Extract the list of search results.
  hits <- dat$response$hits
  
  # If Genius finds no results, return NA.
  if (length(hits) == 0 || nrow(hits) == 0) return(NA_character_)
  
  # Return the first Genius result URL.
  hits$result.url[1]
}

```


## Genius Lyric Scraping

```{r country_genius_searching}
# Function to scrape lyrics from a Genius song URL.
# Input:
#   url = Genius song page URL
# Output:
#   lyrics text, or NA if lyrics cannot be scraped
get_lyrics_from_url <- function(url) {
  
  # If no URL exists, return NA.
  if (is.na(url) || url == "") return(NA_character_)
  
  # Pause randomly between requests so the website is not hit too quickly.
  # This made the full dataset work more reliably than scraping all songs rapidly. This did not seem to be an issue with Rap songs, but it did need to be added for the country songs due to Genius blocking the request.
  Sys.sleep(runif(1, 3, 6))
  
  # Request the Genius webpage.
  # The user agent and headers make the request look more like a normal browser visit.
  res <- tryCatch(
    GET(
      url,
      user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 Chrome/120.0 Safari/537.36"),
      add_headers(
        `Accept-Language` = "en-US,en;q=0.9",
        `Accept` = "text/html,application/xhtml+xml,application/xml;q=0.9,*/*;q=0.8"
      )
    ),
    error = function(e) return(NULL)
  )
  
  # If the webpage request fails, return NA.
  if (is.null(res) || status_code(res) != 200) return(NA_character_)
  
  # Convert the webpage HTML into something R can read and search.
  page <- tryCatch(read_html(res), error = function(e) return(NULL))
  if (is.null(page)) return(NA_character_)
  
  # Extract lyrics from the Genius lyric containers.
  lyrics <- page %>%
    html_elements("div[data-lyrics-container='true']") %>%
    html_text2() %>%
    paste(collapse = "\n")
  
  # If no lyrics are found, return NA.
  if (lyrics == "" || length(lyrics) == 0) return(NA_character_)
  
  # Clean the scraped lyrics by removing the ending "Embed" text
  # and extra spaces.
  lyrics %>%
    str_replace_all("Embed$", "") %>%
    str_squish()
}

```


## Lyric Collection

```{r country_lyric_collection}
## Collect country lyrics and save dataset
# Purpose: Add Genius URLs and lyrics to country songs dataset
# Inputs: country_songs dataframe
# Outputs: country_songs_with_lyrics.csv and progress file
# NOTE: This chunk was run once and should not be re-run when knitting.

country_songs_with_lyrics <- country_songs %>%
  mutate(
    main_artist = str_split(artists, ",|&|feat\\.|featuring") %>% map_chr(1),
    genius_url = map2_chr(main_artist, title, search_genius_url)
  )

# Create an empty vector to store lyrics for each song.
lyrics_list <- vector("character", nrow(country_songs_with_lyrics))

# Loop through each song one at a time.
# This approach is slower than mutate/map, but it is more reliable because:
#   - it pauses between requests
#   - it saves progress after each song, did this because the first few attempts took >20 mins to run and still did not pull the lyrics, so I did not want to loose everything if one or a few songs did not work
for (i in seq_len(nrow(country_songs_with_lyrics))) {
  
  # Print progress in the console so I can see which song is being processed.
  message("Pulling lyrics for song ", i, " of ", nrow(country_songs_with_lyrics), ": ",
          country_songs_with_lyrics$title[i])
  
  # Pull lyrics for the current song using its Genius URL.
  lyrics_list[i] <- get_lyrics_from_url(country_songs_with_lyrics$genius_url[i])
  
  # Add the lyrics collected so far back into the dataset.
  country_songs_with_lyrics$lyrics <- lyrics_list
  
  # Save a progress file after each song.
  # This prevents losing all results if the process stops before finishing.
  write.csv(
    country_songs_with_lyrics,
    "Country_songs_with_lyrics_PROGRESS.csv",
    row.names = FALSE
  )
}

# Save the completed dataset with Genius URLs and lyrics.
write.csv(
  country_songs_with_lyrics,
  "Country_songs_with_lyrics.csv",
  row.names = FALSE
)

# Count how many songs successfully received lyrics.
sum(!is.na(country_songs_with_lyrics$lyrics))

# Count how many songs are still missing lyrics.
sum(is.na(country_songs_with_lyrics$lyrics))

```


## Load Lyrics Dataset

```{r country_load_lyrics}
## Load country lyrics dataset
# Purpose: Load completed country lyrics dataset without re-scraping
# Inputs: Country_songs_with_lyrics.csv
# Outputs: country_songs_with_lyrics dataframe
country_songs <- read_csv("country_songs_with_lyrics.csv")
```


## Word Count Function

```{r country_word_count, message=FALSE, warning=FALSE}
## Clean country songs dataset starting from raw Genius lyrics
## Count words in country songs
# Purpose: Count total words in each song
# Inputs: Lyrics text
# Outputs: Numeric word count

# Load packages needed for reading data, cleaning text, and working with data
library(readr)
library(dplyr)
library(stringr)
library(purrr)

# Read in the dataset that contains the raw scraped lyrics
country_songs_with_lyrics <- read_csv("country_songs_with_lyrics.csv")

# Function to count total number of words in each song
# This will be used later to identify unusually long songs
word_count <- function(lyrics) {
  
  # If lyrics are missing, return NA so we don’t break the code
  if (is.na(lyrics) || lyrics == "") return(NA_real_)
  
  # Convert to lowercase, remove punctuation, and split into words
  words <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_split("\\s+") %>%
    unlist()
  
  # Remove empty values that may have been created
  words <- words[words != ""]
  
  # Return total number of words
  length(words)
}

```


## Lyric Cleaning Function 

```{r country_lyric_cleaning}
## Clean country lyrics
# Purpose: Remove Genius metadata, section labels, and extra text
# Inputs: Raw lyrics text
# Outputs: Cleaned lyrics text

clean_lyrics <- function(lyrics) {
  
  if (is.na(lyrics) || lyrics == "") return(NA_character_)
  
  # Words that usually show up in description text (not lyrics)
  metadata_words <- regex(
    "written|recorded|released|album|single|billboard|chart|peaked|number one|hit number|recounts|chronicles|lead single|second single|song by|song written|music video",
    ignore_case = TRUE
  )
  
  # First round of cleaning
  cleaned <- lyrics %>%
    
    # Remove anything before "Read More" or "Lyrics" (common Genius page markers)
    str_replace("^.*?Read More", "") %>%
    str_replace("^.*?Lyrics", "") %>%
    
    # Remove sentences that start with quotes (usually song descriptions)
    str_replace('^\\s*".*?\\.', "") %>%
    str_replace("^\\s*“.*?\\.", "") %>%
    
    # Remove section labels like [Verse 1], [Chorus], etc.
    str_replace_all("\\[.*?\\]", " ") %>%
    
    # Remove contributor and translation info
    str_replace_all("\\d+ Contributors?", " ") %>%
    str_replace_all("Translations\\s+[A-Za-z ]+", " ") %>%
    
    # Remove leftover "Embed" text from Genius pages
    str_replace_all("\\d*Embed$", " ") %>%
    str_replace_all("Embed$", " ") %>%
    
    # Clean up spacing
    str_squish()
  
  # Second round: remove leftover intro sentences that still look like metadata
  # We check the first sentence a few times and remove it if it looks like description text
  for (i in 1:5) {
    
    first_sentence <- str_extract(cleaned, "^.*?\\.")
    
    if (!is.na(first_sentence) && str_detect(first_sentence, metadata_words)) {
      cleaned <- cleaned %>%
        str_remove("^.*?\\.") %>%
        str_squish()
    }
  }
  
  cleaned
}

```


## Clean Songs Dataset

```{r country_clean_dataset, message=FALSE, warning=FALSE}
## Clean country songs dataset
# Purpose: Apply lyric cleaning and create word count variable
# Inputs: country_songs_with_lyrics dataframe
# Outputs: country_songs_clean dataframe and country_songs_clean.csv

country_songs_clean <- country_songs_with_lyrics %>%
  mutate(
    # Clean the lyrics
    lyrics = map_chr(lyrics, clean_lyrics),
    
    # Count number of words in each song
    number_of_words = map_dbl(lyrics, word_count)
  ) %>%
  
  # Remove rows where lyrics are still missing or empty
  filter(!is.na(lyrics), lyrics != "") %>%
  
  # Remove duplicate song titles so each song is only counted once
  distinct(title, .keep_all = TRUE)

# Check for songs that still look too long (likely still have junk text)
long_songs_check <- country_songs_clean %>%
  filter(number_of_words > 1000) %>%
  select(year, title, artists, number_of_words, lyrics)

# Save the cleaned dataset so we don’t have to redo this step again
write.csv(
  country_songs_clean,
  "country_songs_clean.csv",
  row.names = FALSE
)

# Save the long songs separately so we can inspect them if needed
write.csv(
  long_songs_check,
  "long_songs_check.csv",
  row.names = FALSE
)

# Print number of songs after cleaning
nrow(country_songs_clean)

# Show which songs are still unusually long
long_songs_check %>% select(title, artists, number_of_words)
```


## Stop Words

```{r country_stop_words}
## Load stop words
# Purpose: Load common English stop words to remove from word repetition calculation
# Inputs: tidytext stop_words dataset
# Outputs: stop_words dataframe available for filtering

data("stop_words")

```


## Word Repetition Function

```{r country_word_repetition_function}
## Calculate word repetition score
# Purpose: Measure how repetitive each country song is based on repeated meaningful words
# Inputs: lyrics text from each song
# Outputs: numeric repetition score (higher = more repetition)

word_repetition_score <- function(lyrics) {
  
  # If lyrics are missing or empty, return NA
  # This prevents errors in later calculations
  if (is.na(lyrics) || lyrics == "") return(NA_real_)

  # Convert lyrics to lowercase, remove punctuation, and split into words
  words <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_split("\\s+") %>%
    unlist()

  # Remove empty strings created during splitting
  words <- words[words != ""]

  # Remove common stop words (e.g., "the", "and", "to")
  # These words do not contribute meaningful repetition
  words <- words[!words %in% stop_words$word]

  # If no meaningful words remain, return NA
  if (length(words) == 0) return(NA_real_)

  # Calculate repetition score:
  # 1 - (unique words / total words)
  # Higher score = more repetition
  1 - (length(unique(words)) / length(words))
}
```


## Word Count Function

```{r country_word_count_function}
## Count words in each song
# Purpose: Measure song length using the total number of words in the lyrics
# Inputs: lyrics text from each song
# Outputs: numeric word count

word_count <- function(lyrics) {
  
  # If lyrics are missing or empty, return NA
  # This prevents errors in later calculations
  if (is.na(lyrics) || lyrics == "") return(NA_real_)

  # Convert lyrics to lowercase, remove punctuation, and split into words
  words <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_split("\\s+") %>%
    unlist()

  # Remove empty strings created during splitting
  words <- words[words != ""]

  # Return total number of words in the song
  length(words)
}

```


## Add Word Repetition to Dataset

```{r country_add_word_repetition}
## Add word repetition scores to country dataset
# Purpose: Create word_repeat_score variable
# Inputs: country_songs_clean dataframe
# Outputs: Updated country_songs_clean dataframe

country_songs_clean <- country_songs_clean %>%
  mutate(
    word_repeat_score = map_dbl(lyrics, word_repetition_score)
  )

# Look at results
country_songs_clean %>%
  select(title, word_repeat_score) %>%
  head()
```



## Title Repetition Function

```{r country_title_repetition_function}
## Calculate title repetition score
# Purpose: Count how many times each country song title appears in its lyrics
# Inputs: song title and lyrics text
# Outputs: numeric count of title repetitions

title_repetition_score <- function(title, lyrics) {
  
  # If title or lyrics are missing, return NA
  # This prevents errors during the calculation
  if (is.na(title) || is.na(lyrics) || lyrics == "") return(NA_real_)

  # Clean the song title so capitalization and punctuation do not affect matching
  title_clean <- title %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_squish()

  # Clean the lyrics using the same process as the title
  lyrics_clean <- lyrics %>%
    tolower() %>%
    str_replace_all("[^a-z0-9\\s']", " ") %>%
    str_squish()

  # Count how many times the cleaned title appears in the cleaned lyrics
  str_count(lyrics_clean, fixed(title_clean))
}

```


## Add Title Repetition to Dataset

```{r country_add_title_repetition}
## Add title repetition scores to country dataset
# Purpose: Create title_repeat_score variable
# Inputs: country_songs_clean dataframe
# Outputs: Updated country_songs_clean dataframe

country_songs_clean <- country_songs_clean %>%
  mutate(
    title_repeat_score = map2_dbl(title, lyrics, title_repetition_score)
  )

# View results
country_songs_clean %>%
  select(title, title_repeat_score) %>%
  head()
```


## Yearly Summary

```{r country_yearly_summary}
## Summarize repetition by year
# Purpose: Calculate average repetition scores for each year
# Inputs: country_songs_clean dataframe with repetition score columns
# Outputs: year_summary dataframe for plotting trends over time

year_summary <- country_songs_clean %>%
  group_by(year) %>%
  summarise(
    
    # Average word repetition score per year
    avg_word_repeat = mean(word_repeat_score, na.rm = TRUE),
    
    # Average title repetition count per year
    avg_title_repeat = mean(title_repeat_score, na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  
  # Remove any rows where averages couldn’t be calculated
  filter(!is.na(avg_word_repeat), !is.nan(avg_word_repeat))

# View the summary
year_summary
```


## Word Repetition Plot

```{r country_word_repetition_plot, warning=FALSE}
## Plot word repetition over time
# Purpose: Visualize how average word repetition changes by year for country music

library(ggplot2)

ggplot(year_summary, aes(x = year, y = avg_word_repeat)) +
  
  # Line showing trend over time
  geom_line(linewidth = 1) +
  
  # Points for each year
  geom_point(size = 2) +
  
  # Add a smooth trend line for easier interpretation
  geom_smooth(color = "blue", se = FALSE, linetype = "dashed") +
  
  # Clean theme
  theme_minimal() +
  
  # Improve text readability
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  
  labs(
    title = "Word Repetition in Country Songs Over Time (2006–2026)",
    subtitle = "Average repetition score per year (stopwords removed)",
    x = "Year",
    y = "Repetition Score"
  )

```
The results show a general increase in word repetition in country songs over time. From 2006 to around 2020, repetition scores steadily rise, indicating that more recent songs tend to reuse words more frequently. This suggests a shift toward more repetitive lyrical structures, such as stronger emphasis on choruses or hooks. Although there is some variation across individual years, the overall upward trend is clear. After 2020, repetition appears to level off slightly, which may indicate that repetition has reached a plateau in more recent country music.


## Title Repetition Plot
```{r country_title_repetition_plot, warning=FALSE}
## Plot country title repetition over time
# Purpose: Visualize changes in average title repetition by year
# Inputs: country_year_summary dataframe
# Outputs: Line plot with trend line

ggplot(year_summary, aes(x = year, y = avg_title_repeat)) +
  
  # Line showing trend
  geom_line(linewidth = 1) +
  
  # Points for each year
  geom_point(size = 2) +
  
  # Add smooth trend line
  geom_smooth(color = "blue", se = FALSE, linetype = "dashed") +
  
  # Clean theme
  theme_minimal() +
  
  # Improve readability
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  
  labs(
    title = "Title Repetition in Country Songs Over Time (2006–2026)",
    subtitle = "Average number of times the song title appears in lyrics",
    x = "Year",
    y = "Average Title Repetitions"
  )

```

The title repetition results show a more variable pattern compared to overall word repetition. While there are noticeable spikes in certain years, there is no clear long-term upward or downward trend. This suggests that although country songs may be becoming more repetitive overall, they are not necessarily repeating the song title more frequently. Instead, title repetition appears to fluctuate depending on the style of songs that are popular in a given year.

## Extras

## Top words overall

```{r country_top_words_overall, message=FALSE, warning=FALSE}
## Find most common country words
# Purpose: Identify most frequently used words in country lyrics
# Inputs: country_songs_clean dataframe
# Outputs: country_top_words dataframe

library(tidytext)
library(dplyr)

top_words <- country_songs_clean %>%
  unnest_tokens(word, lyrics) %>%
  anti_join(stop_words, by = "word") %>%
  count(word, sort = TRUE)

# View top words
top_words %>%
  head(20)
```


## Most Repetitive Country Songs

```{r country_most_repetitive}
## Identify most repetitive country songs
# Purpose: Display songs with highest word repetition scores
# Inputs: country_songs_clean dataframe
# Outputs: Table of most repetitive country songs

country_songs_clean %>%
  arrange(desc(word_repeat_score)) %>%
  select(title, artists, word_repeat_score) %>%
  head(10)
```


